---
title: ""
author: "Jane Pascar"
output: pdf_document
fontsize: 12pt
geometry: margin=1in
---
## Jane Pascar  
### HW #4  
  
### Problem 1:
$H_{0}$ : p=0.07, $H_{A}$: p > 0.07  
$\alpha$ = 0.05
```{r echo=FALSE}
# Problem 1
binom.test(x = 4, n = 25, p = .07, alternative = "greater", conf.level = .95)
```
Test statistic: $4/25 = 0.16$  
  
Conclusion: Since $p-value = .09 > \alpha = 0.05$, I cannot reject $H_{0}$.
This means that there is not enough evidence to make the claim that 
the random sample of females have a greater proportion of 
postive test results compared to the proportion reported by the original 
researcher's study. 

### Problem 2:
$H_{0}$ : $p_{1} = p_{2}$, $H_{A}$: $p_{1} \neq p_{2}$  
$\alpha$ = 0.05
```{r echo=FALSE}
rabbits <-
matrix(c(8,5,9,8),
       nrow = 2,
       dimnames = list("2in Arena" = c("Yes", "No"),
                       "1in Arena" = c("Yes", "No")))
rabbits
mcnemar.test(x = rabbits, correct = TRUE)
```
Test statistic: chi-squared = 0.64286
  
Conclusion: Since $p-value = 0.4227 > \alpha = 0.05$, I cannot reject $H_{0}$.
This means that there is not enough evidence to make the claim that the size
of the arena affects the response of rabbit ticks to light.
  
### R Code: 
```{r echo=T, results='hide'}
# Problem 1
binom.test(x = 4, n = 25, p = .07, alternative = "greater", conf.level = .95)

# Problem 2
rabbits <-
matrix(c(8,5,9,8),
       nrow = 2,
       dimnames = list("2in Arena" = c("Yes", "No"),
                       "1in Arena" = c("Yes", "No")))
rabbits
mcnemar.test(x = rabbits, correct = TRUE)
```

