---
title: ""
author: "Jane Pascar"
output: pdf_document
fontsize: 12pt
geometry: margin=1in
---
```{r include=FALSE}

```

## Jane Pascar  
### HW #6  
  
### Problem 1:
$H_{0}$: $F_{1(Women)}=F_{2(Men)}$  
$H_{A}$: $F_{1(Women)}{\neq}F_{2(Men)}$  
$\alpha$ = 0.05  
```{r echo=F}
problem1_data <- data.frame(Women = c(21,37,35,54,31,33,42,27,47,26),
                            Men = c(56,25,43,42,54,33,34,35,47,39))
print(problem1_data)
ks.test(x = problem1_data$Women, y = problem1_data$Men, alternative = "two.sided")
```
D test statistic = 0.3
  
Conclusion: Since the K-S $p-value = 0.7591 > \alpha = 0.05$, I can not reject 
$H_{0}$. This means that the two samples were drawn from identical populations.

### Problem 2:
$H_{0}$: $F_{1(MethodA)}=F_{2(MethodB)}$  
$H_{A}$: $F_{1(MethodB)}{\neq}F_{2(MethodB)}$  
$\alpha$ = 0.05  
```{r echo=F}
problem2_data <- data.frame(A = c(25,22,17,16,19,24,26,21,19,27),
                            B = c(21,22,14,13,22,22,15,21,23,21))
print(problem2_data)
ks.test(x = problem2_data$A, y = problem2_data$B, alternative = "two.sided")
```
D test statistic = 0.4
  
Conclusion: Since the K-S $p-value = 0.4005 > \alpha = 0.05$, I can not reject 
$H_{0}$. This means that the two samples were drawn from identical populations.

### R Code:
```{r echo=T, message=FALSE, warning=FALSE, results='hide', fig.show=FALSE}
problem1_data <- data.frame(Women = c(21,37,35,54,31,33,42,27,47,26),
                            Men = c(56,25,43,42,54,33,34,35,47,39))
print(problem1_data)
ks.test(x = problem1_data$Women, y = problem1_data$Men, alternative = "two.sided")

problem2_data <- data.frame(A = c(25,22,17,16,19,24,26,21,19,27),
                            B = c(21,22,14,13,22,22,15,21,23,21))
print(problem2_data)
ks.test(x = problem2_data$A, y = problem2_data$B, alternative = "two.sided")
```

