---
title: ""
author: "Jane Pascar"
output: pdf_document
fontsize: 12pt
geometry: margin=1in
---
```{r include=FALSE}

```

## Jane Pascar  
### HW #7 
  
### Problem 1:
$H_{0}$ : each correlation coefficient = 0  
$H_{A}$: each correlation coefficient ${\neq}$ 0  
$\alpha$ = 0.05
```{r echo=F}
problem1_data <- data.frame(pecking.order = c(12,5,11,8,4,6,7,1,14,3,10,15,2,9,13),
                            friendliness = c(8,1,9,10,6,5,3,7,11,2,12,13,4,14,15))
plot(x = problem1_data$pecking.order, 
     y = problem1_data$friendliness)
cor(x = problem1_data$pecking.order, 
    y = problem1_data$friendliness, 
    method = "pearson")
cor(x = problem1_data$pecking.order, 
    y = problem1_data$friendliness, 
    method = "spearman")
cor(x = problem1_data$pecking.order, 
    y = problem1_data$friendliness, 
    method = "kendall")
```
Chi-sq test statistic: 35.48 with 1 df
  
Conclusion: Since the Fisher's Exact $p-value = 3.45e-10 << \alpha = 0.05$, I can strongly reject 
$H_{0}$. This means that there is enough evidence to make the claim that 
opportunity level and labor force mobility are associated.  

### Problem 2:
$H_{0}$ : The sampled populations are **homogeneous**  
$H_{A}$: The sampled populations are **not homogeneous**  
$\alpha$ = 0.05  
```{r echo=F}
problem2_data <- matrix(data = c(33,14,23,23), nrow = 2, 
                dimnames = list(c("Solved", "Not Solved"),
                                c("Experimental", "Control")))
print(problem2_data)
chisq.test(x = problem2_data)
fisher.test(x = problem2_data)
```
Conclusion: Since the Fisher's Exact $p-value = 0.0579 > \alpha = 0.05$, I can not reject 
$H_{0}$, although it is only slightly greater than $\alpha$. This means that 
there is not enough evidence to support the claim that the experimental and 
control groups differ with respect to the proportion of solutions.  

### Problem 3:
$H_{0}$ : 
$H_{A}$: 
$\alpha$ = 0.05
```{r echo=F}
problem3_data <- matrix(data = c(3,1,6,3,10,3,1,0,2,2,
                                 19,2,25,15,22,11,6,12,5,11,
                                 39,13,49,41,47,25,14,34,11,37,
                                 14,1,21,35,9,15,5,17,4,8,
                                 10,12,29,26,26,34,11,23,7,20), nrow = 10, 
                dimnames = list("Discipline" = c("Geology", "Biochemistry",
                                                 "Chemisty", "Zoology",
                                                 "Physics", "Engineering",
                                                 "Microbiology", "Botany",
                                                 "Statistics", "Math"),
                                "Funding Category"= c("A", "B", "C", "D", "E")))
print(problem3_data)
chisq.test(x = problem3_data)
fisher.test(x = problem3_data, simulate.p.value = T, B = 100000)

fit <- ca(problem3_data)
fviz_ca_biplot(fit, arrows = c(F,T), title = "")
```

### R Code:
```{r echo=T, message=FALSE, warning=FALSE, results='hide', fig.show=FALSE}
library(ca); library(factoextra)
# Problem 1
problem1_data <- matrix(data = c(45,6,19,43), nrow = 2, 
                dimnames = list("Opportunity" = c("Low", "High"),
                                "Labor Force Mobility" = c("Low", "High")))
chisq.test(x = problem1_data)
fisher.test(x = problem1_data)

# Problem 2
problem2_data <- matrix(data = c(33,14,23,23), nrow = 2, 
                dimnames = list(c("Solved", "Not Solved"),
                                c("Experimental", "Control")))
chisq.test(x = problem2_data)
fisher.test(x = problem2_data)

# Problem 3
problem3_data <- matrix(data = c(3,1,6,3,10,3,1,0,2,2,
                                 19,2,25,15,22,11,6,12,5,11,
                                 39,13,49,41,47,25,14,34,11,37,
                                 14,1,21,35,9,15,5,17,4,8,
                                 10,12,29,26,26,34,11,23,7,20), nrow = 10, 
                dimnames = list("Discipline" = c("Geology", "Biochemistry",
                                                 "Chemisty", "Zoology",
                                                 "Physics", "Engineering",
                                                 "Microbiology", "Botany",
                                                 "Statistics", "Math"),
                                "Funding Category"= c("A", "B", "C", "D", "E")))
chisq.test(x = problem3_data)
fisher.test(x = problem3_data, simulate.p.value = T, B = 100000)

fit <- ca(problem3_data)
fviz_ca_biplot(fit, arrows = c(F,T), title = "")
```

