---
title: ""
author: "Jane Pascar"
output: html_document
fontsize: 12pt
---
```{r include=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
library(BSDA)
library(MultNonParam)
```
## Paired T-Test
#### Example 1
Trace metals in drinking water affect the flavor of the water, and unusually 
high concentrations can pose a health hazard. A researcher reported the 
trace-metals concentrations for both surface water and bottom water in six 
different locations in the South Indian River. Is there any evidence to show 
a difference of metal concentrations between the two layers of water?

```{r echo=FALSE}
trace_metals <- data.frame(Location = c(1:6),
                           Bottom = c(.430,.266,.567,.531,.707,.716),
                           Surface = c(.415,.238,.390,.410,.605,.609),
                           Difference = c(.015,.028,.177,.121,.102,.107))
knitr::kable(trace_metals, 
             col.names = c("Location", "Bottom", "Surface", "$d = Bottom - Surface$"),
             align = "c") %>%
  kable_styling() %>%
  scroll_box(width = "800px", height = "200px")
```
  
Hypotheses:  
$H_{0}: \mu_{d}=0$  
$H_{A}: \mu_{d} \neq 0$  (two-tailed)  
  
$\alpha$ = 0.05  
```{r}
bottom <- c(.430, .266, .567, .531, .707, .716)
surface <- c(.415, .238, .390, .410, .605, .609)

t.test(x = bottom, y = surface, 
       alternative = "two.sided",
       paired = TRUE, # paired by location
       conf.level = .95)
```
Conclusion:  
Since $p-value = 0.014 < \alpha = 0.05$, I have enough evidence to reject the 
$H_{0}$. There is support for the claim that there are different trace metal 
concentrations at the bottom and the surface of the river.  

## One-Sample Sign Test  
#### Example 1  
In a study of transit times, the median transit time in a series of subjects 
was 3.50 seconds. Suppose that another team repeated the procedure on a sample 
of 11 subjects. Could the second team conclude, at the 0.05 level of 
significance, that the median transit time in the population from which its 
sample was drawn is different than 3.50 seconds?
```{r echo=FALSE}
transit_time <- data.frame(Subject = c(1:11),
                           time = c(1.8,3.3,5.65,2.25,2.5,
                                    3.5,2.75,3.25,3.1,2.7,3))
knitr::kable(transit_time, 
             col.names = c("Subject", "Transit Time"),
             align = "c") %>%
  kable_styling() %>%
  scroll_box(width = "800px", height = "200px")
```
  
Hypotheses:  
$H_{0}: M=3.50$  
$H_{A}: M\neq3.5$  (two-tailed)  
  
$\alpha$ = 0.05  
```{r}
transit_time <- c(1.8, 3.3, 5.65, 2.25, 2.5, 3.5, 2.75, 3.25, 3.1, 2.7, 3)
BSDA::SIGN.test(x = transit_time, 
                md = 3.5, 
                alternative = "two.sided", 
                conf.level = .95)
```
Conclusion:  
Since $p-value = 0.02148 < \alpha = 0.05$, I have enough evidence to reject the 
$H_{0}$. There is support for the claim that the population median is not 3.50 
seconds.    
  
#### Example 2
Fifteen heroin addicts were asked to state the age at which they first started 
using the drug. The results are shown in the table below. Can one conclude from 
these data that the median age of the sampled population is 20?  
```{r echo=FALSE}
heroin_age <- data.frame(Subject = c(1:15),
                         Age = c(22,24,37,28,15,14,22,16,18,17,
                                   23,16,20,18,15))
knitr::kable(heroin_age, 
             align = "c") %>%
  kable_styling() %>%
  scroll_box(width = "800px", height = "200px")
```
  
Hypotheses:  
$H_{0}: M=20$  
$H_{A}: M\neq20$  (two-tailed)  
  
$\alpha$ = 0.05  
```{r}
heroin_age <- c(22, 24, 37, 28, 15, 14, 22, 16, 18, 17, 23, 16, 20, 18, 15)
BSDA::SIGN.test(x = heroin_age, 
                md = 20, 
                alternative = "two.sided", 
                conf.level = .95)
```
Conclusion:  
Since $p-value = 0.7905 > \alpha = 0.05$, I have do not have enough evidence to 
reject the $H_{0}$. The median age of heroin addicts is not 20.    
  
## One-Sample Wilcoxon Signed-Rank Test
#### Example 1
In a study of drug abuse in a suburban area, investigators found that the 
median IQ of arrested abusers who were 16 years of age or older was 107. 
Suppose that a researcher wishes to know whether to conclude that the median IQ 
of arrested abusers who are 16 or older in another suburban area is different 
from 107. The following table shows the IQs of a random sample of 15 persons 
from the population of interest. What can the researcher conclude?
```{r echo=FALSE}
drug_iq <- data.frame(Subject = c(1:15),
                         IQ = c(99,100,90,94,135,108,107,111,119,104,127,109,
                                117,105,125))
knitr::kable(drug_iq, 
             align = "c") %>%
  kable_styling() %>%
  scroll_box(width = "800px", height = "200px")
```
  
Hypotheses:  
$H_{0}: M=107$  
$H_{A}: M\neq107$  (two-tailed)  
  
$\alpha$ = 0.05  
```{r}
drug_iq <- c(99, 100, 90, 94, 135, 108, 107, 111, 
             119, 104, 127, 109, 117, 105, 125)
wilcox.test(x = drug_iq, 
            alternative = "two.sided", 
            mu = 107,
            conf.level = .05,
            exact = F)
```
Conclusion:  
Since $p-value = 0.4702 > \alpha = 0.05$, I have do not have enough evidence to 
reject the $H_{0}$. The median IQ of arrested abusers is not 107.  
  
#### Example 2
A test to measure knowledge of current events was given to a sample of 25 
elementary school children from an inner-city neighborhood. The scores are in 
the following table. Can one conclude from these data that the population 
median is 70? 
```{r echo=FALSE}
children_scores <- data.frame(Subject = c(1:25),
                              Scores = c(80,68,30,67,70,62,69,65,53,29,65,68,62,
                                         56,46,48,39,72,36,69,40,61,54,53,25))
knitr::kable(children_scores, 
             align = "c") %>%
  kable_styling() %>%
  scroll_box(width = "800px", height = "200px")
```
  
Hypotheses:  
$H_{0}: M=70$  
$H_{A}: M\neq70$ (two-tailed)  
  
$\alpha$ = 0.05  
```{r}
children_scores <- c(80, 68, 30, 67, 70, 62, 69, 65, 53, 29, 65, 68, 62, 
                     56, 46, 48, 39, 72, 36, 69, 40, 61, 54, 53, 25)
wilcox.test(x = children_scores, 
            alternative = "two.sided", 
            mu = 70,
            conf.level = .05,
            exact = F)
```
Conclusion:  
Since $p-value = 0.0001357 < \alpha = 0.05$, I have enough evidence to 
reject the $H_{0}$. The median score of elementary school children on the 
current-events test is not 70.  
  
## Two-Sample Median Test  
#### Example 1
The stroke-index values were measured for patients admitted to a university 
hospital. We wish to know whether these data provide sufficient evidence to 
indicate that the medians of the two populations represented by the sample 
data are different.  
```{r echo=FALSE}
stroke_index <- data.frame(samp1 = c(25,25,17,26,18,30,24,21,13,30,20,23,26,12,
                                     20,37,9,17,37,20,11,32,16,31,46,20,25,17,
                                     36,54,8,26),
                           samp2 = c(31,21,38,19,38,41,68,28,43,42,30,20,29,13,
                                     32,30, rep(NA, 16)))
options(knitr.kable.NA = '')
knitr::kable(stroke_index, 
             col.names = c("Sample 1", "Sample 2"),
             align = "c") %>%
  kable_styling() %>%
  scroll_box(width = "800px", height = "200px")
```
  
Hypotheses:  
$H_{0}: M_{sample1} = M_{sample2}$  
$H_{A}: M_{sample1} \neq M_{sample2}$ (two-tailed)  
  
$\alpha$ = 0.05  
```{r}
samp1 = c(25, 25, 17, 26, 18, 30, 24, 21, 13, 30, 20, 23, 26, 12, 20, 37, 
          9, 17, 37, 20, 11, 32, 16, 31, 46, 20, 25, 17, 36, 54, 8, 26)
samp2 = c(31, 21, 38, 19, 38, 41, 68, 28, 43, 42, 30, 20, 29, 13, 32, 30)
MultNonParam::mood.median.test(x = samp1, y = samp2, exact = TRUE)
```
Conclusion:  
Since $p-value = 0.0305 < \alpha = 0.05$, I have enough evidence to 
reject the $H_{0}$. The two populations of admitted patients at each of the two 
hospitals do not have the same median stroke-index values.  
  
## Mann-Whitney-Wilcoxon Test  
#### Example 1
A study reported the results of an attempt to assess the predictive validity 
of Klopfer's Prognostic Rating Scale (PRS) with subjects who received behavior 
modification psychotherapy. Following psychotherapy, the subjects were 
separtated into two groups: improved and unimproved. We wish to see whether we 
can conclude on the basis of these data that the two represented populations 
are different with respect to their medians. 
```{r echo=FALSE}
prs_data <- data.frame(Improved = c(2.2,4.1,4.2,5.0,6.3,6.8,6.9,7.1,7.4,7.4,7.7,
                                    8.2,8.7,9.4,9.5,11.7,11.9),
                       Unimproved = c(1.7,2.4,3.3,3.9,4.3,5.0,5.1,5.4,
                                      5.8,6.6, rep(NA,7)))
options(knitr.kable.NA = '')
knitr::kable(prs_data, #
             align = "c") %>%
  kable_styling() %>%
  scroll_box(width = "800px", height = "200px")
```
  
Hypotheses:  
$H_{0}: M_{Improved} = M_{Unimproved}$  
$H_{A}: M_{Improved} \neq M_{Unimproved}$ (two-tailed)  
  
$\alpha$ = 0.05  
```{r}
Improved = c(2.2, 4.1, 4.2, 5.0, 6.3, 6.8, 6.9, 7.1, 7.4, 7.4, 7.7, 
             8.2, 8.7, 9.4, 9.5, 11.7, 11.9)
Unimproved = c(1.7, 2.4, 3.3, 3.9, 4.3, 5.0, 5.1, 5.4, 5.8, 6.6)
wilcox.test(x = Improved, y = Unimproved, 
            alternative = "two.sided", 
            paired = FALSE, 
            exact = FALSE, 
            correct = TRUE, 
            conf.level = .95)
```
Conclusion:  
Since $p-value = 0.003579 < \alpha = 0.05$, I have enough evidence to 
reject the $H_{0}$. The two improved and unimproved groups do not have the same 
median PRS scores.  
  
## Paired Two-Sample Sign Test  
#### Example 1  
A study investigated the effects of togetherness on heart rate in rats. They 
recorded the heart rate of 10 rats while they were alone and while in the 
presence of another rat. The researchers wanted to test if the togetherness in 
rats increases heart rate. 
